---
title: "Analysis of the Results from 'Monitoring global development aid with machine learning'"
author: "Connor Flynn, Quinn McHenry, Katherine Gonzales"
date: "5/5/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(plotly)
library(readxl)
library(janitor)
library(sunburstR)
library(htmltools)
options(scipen = 99)
```


#!#[]#(/Users/connorflynn/Documents/SDG/SIDS_SDG_UNSTATS/visualizations/cifal_logo.png){width=150} 


The data source of this work is the results from the paper [Monitoring global development aid with machine learning](https://www.nature.com/articles/s41893-022-00874-z) specifically sourced form their [Github](https://github.com/MalteToetzke/Monitoring-Global-Development-Aid-With-Machine-Learning/tree/main/Data)


### Goal of this work

Visualize **more granular descriptions of aid than the 17 SDG's** for Small Island Develping States (SIDS)

  - How much aid is each SIDS receiving?:
    - For which sectors? (31 total)
    
  - For SIDS as a whole, what proportion of the overall aid goes to each of the specific topics? (171 total)






```{r, include=FALSE, results='hide'}
activity_clusters <- read_csv(here("data/activity_cluster_data.csv"))
```




```{r, include=FALSE, results='hide'}
descriptive_stats <- read_excel(here("data/descriptive_statistics_clusters.xlsx"))
```



```{r, include=FALSE, results='hide'}
activity_clusters <- activity_clusters %>% 
  mutate(disbursement = USD_Disbursement * 1000000)
```





```{r, include=FALSE, results='hide'}
total_disbursments_by_country <- activity_clusters %>% 
  group_by(RecipientName) %>% 
  summarise(disbursement = sum(disbursement))
```




```{r, include=FALSE, results='hide'}
total_disbursments_by_country_sector <-activity_clusters %>% 
  group_by(RecipientName, Sector) %>% 
  summarise(disbursement = sum(disbursement))
```



```{r, include=FALSE, results='hide'}
unique(activity_clusters$RecipientName)
```



```{r, include=FALSE, results='hide'}
sids <- c('Anguilla', 'Antigua and Barbuda', 'Bahrain', 'Barbados', 'Belize', 'Cabo Verde', 'Comoros', 
             'Cook Islands', 'Dominica', 'Dominican Republic', 'Fiji', 'Grenada', 'Guinea-Bissau', 'Guyana', 
             'Haiti', 'Jamaica', 'Kiribati', 'Maldives', 'Marshall Islands', 'Mauritius', 'Micronesia', 
             'Montserrat', 'Nauru', 'Niue', 'Palau', 'Papua New Guinea', 'Saint Kitts and Nevis', 'Saint Lucia', 
             'Saint Vincent and the Grenadines', 'Samoa', 'Sao Tome and Principe', 'Seychelles', 'Solomon Islands', 
             'Suriname', 'Timor-Leste', 'Tokelau', 'Tonga', 'Trinidad and Tobago', 'Turks and Caicos Islands', 
             'Tuvalu', 'Vanuatu')
```




```{r, include=FALSE, results='hide'}
activity_clusters_sids <- activity_clusters %>%
  filter(RecipientName %in% sids)
```



```{r, include=FALSE, results='hide'}
unique(activity_clusters_sids$RecipientName)
```



```{r, include=FALSE, results='hide'}
total_disbursments_by_sids <- activity_clusters_sids %>% 
  group_by(RecipientName) %>% 
  summarise(disbursement = sum(disbursement))
```



```{r, include=FALSE, results='hide'}
total_disbursments_by_sids_sector <- activity_clusters_sids %>% 
  group_by(RecipientName, Sector) %>% 
  summarise(disbursement = sum(disbursement))
```



```{r, include=FALSE, results='hide'}
plot <- ggplot(total_disbursments_by_sids_sector, aes(x = disbursement,
                                              y = reorder(RecipientName, disbursement),
                                              fill = Sector)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Aid by SIDS by Sector", y = "SIDS", x = "Aid in $USD") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, size = 4, hjust = 1))

plot
```

**_Below is a visualization of each SIDS and their total disbursement (funding), color coordinated to illustrate what fraction of the budget was used for a given sector. The results indicate that of the 31 SIDS, Haiti received the most funds, with the majority directed towards the 'Emergency Response' sector. Hovering the cursor over a section of a bar will display the disbursement and sector for that section._**

```{r}
ggplotly(plot)
```




```{r, include=FALSE, results='hide'}
total_disbursments_by_sids_sector <- total_disbursments_by_sids_sector %>% 
  group_by(RecipientName)  %>% 
  mutate(percentage = disbursement/sum(disbursement))

```



```{r, include=FALSE, results='hide'}
tile_plot <- ggplot(total_disbursments_by_sids_sector, aes(x=Sector , y= RecipientName,
                                              fill = percentage)) +
  geom_tile() +
  labs(title = "Percentage of Budget allocated to each Sector by SIDS", y = "SIDS", x = "Sector") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, size = 4, hjust = 1),
        axis.text.y = element_text(size = 3, hjust = 1)) +
  scale_fill_gradient(low = "khaki", high = "red")
  
tile_plot

```

**_Next, we have a heat-map of the distribution of funds by sector. Each tile represents the percentage of that SIDS' total budget that was spent on the corresponding sector. Darker colors indicate a higher percentage; for example, Bahrain used over 70% of its' funding on Education, indicated by the dark red tile on the lower left side of the map. Hovering the cursor over a given tile will display the sector, SIDS, and percentage distribution - expressed as a decimal - for that tile._**
```{r}
ggplotly(tile_plot)
```






```{r, include=FALSE, results='hide'}
colnames(activity_clusters)[colnames(activity_clusters) == "cluster"] ="Cluster"
```



```{r, include=FALSE, results='hide'}
activity_stats_join<- full_join(activity_clusters, descriptive_stats, by='Cluster')
```




```{r, include=FALSE, results='hide'}
activity_stats_sids<-activity_stats_join %>%
  filter(RecipientName %in% sids)
```




```{r, include=FALSE, results='hide'}
activity_stats_sids <- activity_stats_sids %>% 
  clean_names()
```


```{r, include=FALSE, results='hide'}
topic_sector_sunburst<- activity_stats_sids %>% 
  group_by(topic, sector) %>% 
  summarize(sum_disbursement = sum(disbursement))
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_merged <- topic_sector_sunburst %>%
  filter(sector != "") %>%
  mutate(path = paste(sector, topic, sep="-")) %>%
  dplyr::select(path, sum_disbursement)
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_merged<- subset(topic_sector_sunburst_merged, select= -c(topic))
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_merged$path <- gsub(" - ", ",", topic_sector_sunburst_merged$path)
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_merged$path <- gsub("^(.+)-(.+)-", "\\1 \\2-", topic_sector_sunburst_merged$path)
```

**_This sunburst chart allows us to better visualize the relationship between the SIDS sectors and their overarching "topics" as a collective, representing the data for all disbursement of all SIDS, combined. The inner ring represents the proportion of funding that was distributed to each topic, of which there are 31. The outer ring shows us the distribution of funding for each sector, separated by their respective topics. Hovering the cursor over a section of the chart will display the proportion expressed as a percentage, along with labels in the upper left that indicate the relative topic and/or sector._**

```{r}
p <- sunburst(topic_sector_sunburst_merged, legend=FALSE, width="100%", height=400)
p
```

```{r, include=FALSE, results='hide'}
AIS<-c('Bahrain', 'Cabo Verde', 'Comoros', 'Guinea-Bissau', 'Maldives', 'Mauritius',' Sao Tomé and Principe', 'Seychelles', 'Singapore')
```


```{r, include=FALSE, results='hide'}
Caribbean<-c('Antigua and Barbuda', 'Bahamas', 'Barbados', 'Belize', 'Cuba', 'Dominica', 'Dominican Republic', 'Grenada', 'Guyana', 'Haiti', 'Jamaica', 'Saint Kitts and Nevis', 'Saint Lucia', 'Saint Vincent and the Grenadines', 'Suriname', 'Trinidad and Tobago', 'Turks and Caicos Islands', 'Montserrat', 'Anguilla')
```

```{r, include=FALSE, results='hide'}
Pacific<-c('Fiji', 'Kiribati', 'Marshall Islands', 'Micronesia (Federated States of)', 'Nauru', 'Palau', 'Papua New Guinea', 'Samoa', 'Solomon Islands', 'Timor-Leste', 'Tonga', 'Tuvalu', 'Vanuatu')
```




```{r, include=FALSE, results='hide'}
activity_stats_pacific<-activity_stats_join %>%
  filter(RecipientName %in% Pacific)
```

```{r, include=FALSE, results='hide'}
activity_stats_pacific <- activity_stats_pacific %>% 
  clean_names()
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_pacific<- activity_stats_pacific %>% 
  group_by(topic, sector) %>% 
  summarize(sum_disbursement = sum(disbursement))
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_pacific_merged <- topic_sector_sunburst_pacific %>%
  filter(sector != "") %>%
  mutate(path = paste(sector, topic, sep="-")) %>%
  dplyr::select(path, sum_disbursement)
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_pacific_merged<- subset(topic_sector_sunburst_pacific_merged, select= -c(topic))
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_pacific_merged$path <- gsub(" - ", ",", topic_sector_sunburst_pacific_merged$path)
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_pacific_merged$path <- gsub("^(.+)-(.+)-", "\\1 \\2-", topic_sector_sunburst_pacific_merged$path)
```

```{r, include=FALSE, results='hide'}
pac <- sunburst(topic_sector_sunburst_pacific_merged, legend=FALSE, width="100%", height=400)
pac
```




```{r, include=FALSE, results='hide'}
activity_stats_caribbean<-activity_stats_join %>%
  filter(RecipientName %in% Caribbean)
```

```{r, include=FALSE, results='hide'}
activity_stats_caribbean <- activity_stats_caribbean %>% 
  clean_names()
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_caribbean<- activity_stats_caribbean %>% 
  group_by(topic, sector) %>% 
  summarize(sum_disbursement = sum(disbursement))
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_caribbean_merged <- topic_sector_sunburst_caribbean %>%
  filter(sector != "") %>%
  mutate(path = paste(sector, topic, sep="-")) %>%
  dplyr::select(path, sum_disbursement)
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_caribbean_merged<- subset(topic_sector_sunburst_caribbean_merged, select= -c(topic))
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_caribbean_merged$path <- gsub(" - ", ",", topic_sector_sunburst_caribbean_merged$path)
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_caribbean_merged$path <- gsub("^(.+)-(.+)-", "\\1 \\2-", topic_sector_sunburst_caribbean_merged$path)
```

```{r, include=FALSE, results='hide'}
car <- sunburst(topic_sector_sunburst_caribbean_merged, legend=FALSE, width="100%", height=400)
car
```


```{r, include=FALSE, results='hide'}
activity_stats_AIS<-activity_stats_join %>%
  filter(RecipientName %in% AIS)
```

```{r, include=FALSE, results='hide'}
activity_stats_AIS <- activity_stats_AIS %>% 
  clean_names()
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_AIS<- activity_stats_AIS %>% 
  group_by(topic, sector) %>% 
  summarize(sum_disbursement = sum(disbursement))
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_AIS_merged <- topic_sector_sunburst_AIS %>%
  filter(sector != "") %>%
  mutate(path = paste(sector, topic, sep="-")) %>%
  dplyr::select(path, sum_disbursement)
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_AIS_merged<- subset(topic_sector_sunburst_AIS_merged, select= -c(topic))
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_AIS_merged$path <- gsub(" - ", ",", topic_sector_sunburst_AIS_merged$path)
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_AIS_merged$path <- gsub("^(.+)-(.+)-", "\\1 \\2-", topic_sector_sunburst_AIS_merged$path)
```

```{r, include=FALSE, results='hide'}
AIS <- sunburst(topic_sector_sunburst_AIS_merged, legend=FALSE, width="100%", height=400)
AIS
```

**_Finally, we have three different sunbursts that illustrate the budgets of SIDS based on their region. From left to right; the Caribbean region, the Pacific region, and the Atlantic/Indian Ocean/South China Sea region. Compare these charts with the original sunburst to better understand how the funding is disbursed regionally. For a list of which SIDS belong to which region, visit https://sustainabledevelopment.un.org/topics/sids/list. _**

```{r}
div(
  style="display: flex; align-items:center;",
  div(style="width:50%; border:1px solid #ccc;", car),
  div(style="width:50%; border:1px solid #ccc;", pac),
  div(style="width:50%; border:1px solid #ccc;", AIS)
)

```


