---
title: "Sunburst by Sids"
author: "Connor"
date: "5/12/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(plotly)
library(readxl)
library(janitor)
library(sunburstR)
options(scipen = 99)
```


```{r, include=FALSE, results='hide'}
activity_clusters <- read_csv(here("data/activity_cluster_data.csv"))
```




```{r, include=FALSE, results='hide'}
descriptive_stats <- read_excel(here("data/descriptive_statistics_clusters.xlsx"))
```



```{r, include=FALSE, results='hide'}
activity_clusters <- activity_clusters %>% 
  mutate(disbursement = USD_Disbursement * 1000000)
```





```{r, include=FALSE, results='hide'}
colnames(activity_clusters)[colnames(activity_clusters) == "cluster"] ="Cluster"
```



```{r, include=FALSE, results='hide'}
activity_stats_join<- full_join(activity_clusters, descriptive_stats, by='Cluster')
```


```{r, include=FALSE, results='hide'}
sids <- c('Anguilla', 'Antigua and Barbuda', 'Bahrain', 'Barbados', 'Belize', 'Cabo Verde', 'Comoros', 
             'Cook Islands', 'Dominica', 'Dominican Republic', 'Fiji', 'Grenada', 'Guinea-Bissau', 'Guyana', 
             'Haiti', 'Jamaica', 'Kiribati', 'Maldives', 'Marshall Islands', 'Mauritius', 'Micronesia', 
             'Montserrat', 'Nauru', 'Niue', 'Palau', 'Papua New Guinea', 'Saint Kitts and Nevis', 'Saint Lucia', 
             'Saint Vincent and the Grenadines', 'Samoa', 'Sao Tome and Principe', 'Seychelles', 'Solomon Islands', 
             'Suriname', 'Timor-Leste', 'Tokelau', 'Tonga', 'Trinidad and Tobago', 'Turks and Caicos Islands', 
             'Tuvalu', 'Vanuatu')
```

```{r, include=FALSE, results='hide'}
activity_stats_sids<-activity_stats_join %>%
  filter(RecipientName %in% sids)
```




```{r, include=FALSE, results='hide'}
activity_stats_sids <- activity_stats_sids %>% 
  clean_names()
```


```{r, include=FALSE, results='hide'}
topic_sector_sunburst<- activity_stats_sids %>% 
  group_by(topic, sector) %>% 
  summarize(sum_disbursement = sum(sum_disbursement))
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_merged <- topic_sector_sunburst %>%
  filter(sector != "") %>%
  mutate(path = paste(sector, topic, sep="-")) %>%
  dplyr::select(path, sum_disbursement)
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_merged<- subset(topic_sector_sunburst_merged, select= -c(topic))
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_merged$path <- gsub(" - ", ",", topic_sector_sunburst_merged$path)
```

```{r, include=FALSE, results='hide'}
topic_sector_sunburst_merged$path <- gsub("^(.+)-(.+)-", "\\1 \\2-", topic_sector_sunburst_merged$path)
```


```{r}
p <- sunburst(topic_sector_sunburst_merged, legend=TRUE)
p
```



Sunburst by SIDS - All SIDS

```{r}
activity_stats_sids_by_sids<- activity_stats_sids %>% 
  group_by(topic, sector, recipient_name) %>% 
  summarize(sum_disbursement = sum(sum_disbursement))
  
```


```{r}
topic_sector_sunburst_merged_by_sids <- activity_stats_sids_by_sids %>%
  filter(sector != "") %>%
  mutate(path = paste(sector, topic, sep="-")) %>%
  dplyr::select(path, sum_disbursement, recipient_name)
```

Chat GPT Notes

# install and load the required packages
install.packages("sunburstR")
library(sunburstR)

# load the data
sunburst_sids <- read.csv("your_file.csv") # replace "your_file.csv" with the name of your data file

# create the sunburst plot with all recipient names
sunburst(sunburst_sids, 
         count = "amount_usd",
         legend = "recipient_name")

# add the interactive dropdown for recipient names
update_sunburst_input(
  input_id = "recipient_name_dropdown",
  data = sunburst_sids,
  label = "Select recipient name:",
  column = "recipient_name",
  type = "select",
  multiple = FALSE
)

# create a reactive function to update the sunburst plot based on the selected recipient name
ui <- fluidPage(
  selectInput(
    inputId = "recipient_name_dropdown",
    label = "Select recipient name:",
    choices = unique(sunburst_sids$recipient_name)
  ),
  sunburstOutput("sunburst_plot")
)

server <- function(input, output) {
  output$sunburst_plot <- renderSunburst({
    selected_data <- sunburst_sids[sunburst_sids$recipient_name == input$recipient_name_dropdown, ]
    sunburst(selected_data,
             count = "amount_usd",
             legend = "recipient_name")
  })
}

# run the app
shinyApp(ui, server)

