---
title: "Solar | Aid & Indicators"
author: "Connor Flynn"
date: "9/14/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(plotly)
library(readxl)
library(janitor)
library(sunburstR)
library(htmltools)
library(dplyr)
library(shiny)
library(tseries)
library(astsa)
options(scipen = 99)
```


Import Data from "Monitoring GLobal Aid with Machine Learning"

```{r, include=FALSE, results='hide'}
activity_clusters <- read_csv(here("data/activity_cluster_data.csv"))
```


```{r, include=FALSE, results='hide'}
descriptive_stats <- read_excel(here("data/descriptive_statistics_clusters.xlsx"))
```

Import Data from SIDS Data Platform - *may normalize for population*

```{r}
installed_renewable_per_capita <- read_csv(here("energy/indicator/installed_renewable_electricity_generating_capacity_watts_per_capita.csv"), skip = 1)

### IMmporting data from file path
energy_intensity_level_of_primary_energy <- read_csv(here("energy/indicator/energy_intensity_level_of_primary_energy(mejajoules_per_constant_2017_purchasisn_power_parity_GDP).csv"), skip = 1)

### IMmporting data from file path
Proportion_of_population_with_access_to_electricity <- read_csv(here("energy/indicator/Proportion_of_population_with_access_to_electricity_by_urban%.csv"), skip = 1)

international_fiancial_flows <- read_csv(here("energy/indicator/international_financial_flows.csv"), skip = 1)

```




```{r}
installed_renewable_per_capita[installed_renewable_per_capita == 'No Data'] <- NA

### New code
energy_intensity_level_of_primary_energy[energy_intensity_level_of_primary_energy == 'No Data'] <- NA

### New code
Proportion_of_population_with_access_to_electricity[Proportion_of_population_with_access_to_electricity == 'No Data'] <- NA

### New Code
international_fiancial_flows[international_fiancial_flows == 'No Data'] <- NA

```




```{r}
installed_renewable_per_capita_longer <- installed_renewable_per_capita %>% 
  mutate(across(!Year, as.character)) %>%
  pivot_longer(cols = !Year,
               names_to = "recipient_name",
               values_to = "value") %>% 
  clean_names()
```

```{r}
### Andrews implementation
energy_intensity_level_of_primary_energy <- energy_intensity_level_of_primary_energy %>% 
  mutate(across(!Year, as.character)) %>%
  pivot_longer(cols = !Year,
               names_to = "recipient_name",
               values_to = "value") %>% 
  clean_names()

Proportion_of_population_with_access_to_electricity <- Proportion_of_population_with_access_to_electricity %>% 
  mutate(across(!Year, as.character)) %>%
  pivot_longer(cols = !Year,
               names_to = "recipient_name",
               values_to = "value") %>% 
  clean_names()


international_fiancial_flows <- international_fiancial_flows %>% 
  mutate(across(!Year, as.character)) %>%
  pivot_longer(cols = !Year,
               names_to = "recipient_name",
               values_to = "value") %>% 
  clean_names()
```



Filter and Join Data from "Monitoring Global Aid with Machine Learning"




```{r, include=FALSE, results='hide'}
activity_clusters <- activity_clusters %>% 
  mutate(disbursement = USD_Disbursement * 1000000)
```

```{r, include=FALSE, results='hide'}
sids <- c('Anguilla', 'Antigua and Barbuda', 'Bahrain', 'Barbados', 'Belize', 'Cabo Verde', 'Comoros', 
             'Cook Islands', 'Dominica', 'Dominican Republic', 'Fiji', 'Grenada', 'Guinea-Bissau', 'Guyana', 
             'Haiti', 'Jamaica', 'Kiribati', 'Maldives', 'Marshall Islands', 'Mauritius', 'Micronesia', 
             'Montserrat', 'Nauru', 'Niue', 'Palau', 'Papua New Guinea', 'Saint Kitts and Nevis', 'Saint Lucia', 
             'Saint Vincent and the Grenadines', 'Samoa', 'Sao Tome and Principe', 'Seychelles', 'Solomon Islands', 
             'Suriname', 'Timor-Leste', 'Tokelau', 'Tonga', 'Trinidad and Tobago', 'Turks and Caicos Islands', 
             'Tuvalu', 'Vanuatu')
```

```{r, include=FALSE, results='hide'}
colnames(activity_clusters)[colnames(activity_clusters) == "cluster"] ="Cluster"
```

```{r, include=FALSE, results='hide'}
activity_stats_join<- full_join(activity_clusters, descriptive_stats, by='Cluster')
```

```{r, include=FALSE, results='hide'}
activity_stats_sids<-activity_stats_join %>%
  filter(RecipientName %in% sids)
```

```{r, include=FALSE, results='hide'}
activity_stats_sids <- activity_stats_sids %>% 
  clean_names()
```


Combining
```{r}
activity_stats_sids_energy <- activity_stats_sids %>% 
  filter(sector == "Energy")
```


```{r}
colnames(installed_renewable_per_capita_longer)[colnames(installed_renewable_per_capita_longer) == "value"] ="Renewable_watts_per_capita"
```


```{r}
indicator_installed_renewable_aid <- activity_stats_sids_energy %>%
  left_join(installed_renewable_per_capita_longer, 
            by = c("year", "recipient_name"))
```




```{r}
colnames(energy_intensity_level_of_primary_energy)[colnames(energy_intensity_level_of_primary_energy) == "value"] ="Energy_intensity_level_of_primary_energy_mj_per_gdp"


colnames(Proportion_of_population_with_access_to_electricity)[colnames(Proportion_of_population_with_access_to_electricity) == "value"] ="Proportion_of_population_with_access_to_electricity"

colnames(international_fiancial_flows)[colnames(international_fiancial_flows) == "value"] ="international_financial_flow"
```

```{r}
indicator_installed_renewable_energy_intensity_aid <- indicator_installed_renewable_aid %>%
  left_join(energy_intensity_level_of_primary_energy, 
            by = c("year", "recipient_name"))
```

```{r}

indicator_installed_renewable_energy_intensity_population_access_aid <- indicator_installed_renewable_energy_intensity_aid %>%
  left_join(Proportion_of_population_with_access_to_electricity, 
            by = c("year", "recipient_name"))
```

```{r}

indicator_installed_renewable_energy_intensity_population_access_financial_flows_aid <- indicator_installed_renewable_energy_intensity_population_access_aid %>%
  left_join(international_fiancial_flows, 
            by = c("year", "recipient_name"))
```


```{r}
energy_aid_indicators <- indicator_installed_renewable_energy_intensity_population_access_financial_flows_aid
```


```{r}
energy_aid_indicators <- energy_aid_indicators %>% 
  select(-cluster, -usd_disbursement, -country_code, -incomegroup_name, -matching_purpose_code, -purpose_code, -count_activities)
```


```{r}
energy_aid_indicators$Renewable_watts_per_capita <- as.numeric(energy_aid_indicators$Renewable_watts_per_capita)
```


*Analysis*

```{r}
unique(energy_aid_indicators$topic)
```
```{r}
renewable_watts_relevant_topics <- c("Energy generation, renewable sources - multiple technologies", "Hydro-electric power plants", "Natural gas-fired electric power plants", "Solar energy", "Geothermal energy", "Marine energy", "Wind energy", "Biofuel-fired power plants")
```



```{r}
energy_aid_indicators_renewable_watts_relevant_topics <- energy_aid_indicators %>% 
  filter(topic %in% renewable_watts_relevant_topics)
```

```{r}
energy_aid_indicators_renewable_watts_relevant_topics_grouped <- energy_aid_indicators_renewable_watts_relevant_topics %>% 
  group_by(year, topic) %>% 
  summarise(topic_disbursement = sum(disbursement)) 



  ggplot(data = energy_aid_indicators_renewable_watts_relevant_topics_grouped, aes(x = year, y = topic_disbursement, color = topic)) +
    geom_line() +
  theme_minimal()
  
              
```

```{r}
energy_aid_indicators_renewable_watts_relevant_topics_grouped <- energy_aid_indicators_renewable_watts_relevant_topics %>% 
  group_by(recipient_name, year, Renewable_watts_per_capita) %>% 
  summarize(renewable_disbursement = sum(disbursement))
```



By Country Correlation - need to add lags
```{r}


# Calculate correlation between disbursement and Renewable_watts_per_capita for each country
correlations <- by(energy_aid_indicators_renewable_watts_relevant_topics_grouped, energy_aid_indicators_renewable_watts_relevant_topics_grouped$recipient_name, function(sub_data) {
  if (all(!is.na(sub_data$renewable_disbursement)) && all(!is.na(sub_data$Renewable_watts_per_capita))) {
    cor(sub_data$renewable_disbursement, sub_data$Renewable_watts_per_capita)
  } else {
    NA
  }
})

# Print the correlations for each country
print(correlations)



```
```{r}
energy_aid_indicators_renewable_watts_relevant_topics_grouped_country <- energy_aid_indicators_renewable_watts_relevant_topics_grouped %>% 
  filter(recipient_name == "Anguilla")  
  
ggplot(energy_aid_indicators_renewable_watts_relevant_topics_grouped_country, aes(x = year, y = renewable_disbursement)) +
  geom_line()

ggplot(energy_aid_indicators_renewable_watts_relevant_topics_grouped_country, aes(x = year, y = Renewable_watts_per_capita)) +
  geom_line()
  
```


Average SIDS Renewable watts per capita per year with renewable disbursement (ccf)


Does the most change in Renewable Watts Per Capita mean the most money

```{r}
all_sids_aid_watts <- energy_aid_indicators_renewable_watts_relevant_topics_grouped %>% 
  group_by(year) %>% 
  summarise(sum_aid = sum(renewable_disbursement),
            sum_watts = sum(Renewable_watts_per_capita, na.rm = TRUE))

  
```



```{r}
ggplot(all_sids_aid_watts, aes(x = year, y = sum_aid)) +
  geom_line()

ggplot(all_sids_aid_watts, aes(x = year, y = sum_watts)) +
  geom_line()
```
```{r}
watts_aid_plot <- ggplot(all_sids_aid_watts, aes(x = sum_aid, y = sum_watts, label = year)) +
  geom_point() +
  geom_smooth()
ggplotly(watts_aid_plot)

lm(sum_watts ~ sum_aid, all_sids_aid_watts)
```



For Andrew:

Are SIDS who get the most funding for coal and oil topics the most stagnant in renewable watts per capita over time

Can we add temporal lag analysis (ccf package/function form video) for 
- All SIDS
- Each SIDS

```{r}
indices_to_extract <- c("Oil-fired electric power plants")

indices_to_extract2 <- c("Coal-fired electric power plants")

```

```{r}
oil_funding <- subset(energy_aid_indicators, topic %in% indices_to_extract)

coal_funding <- subset(energy_aid_indicators, topic %in% indices_to_extract2)
```


```{r}
print(oil_funding)

```





```{r}
ccf_result <- ccf(oil_funding$disbursement, oil_funding$Renewable_watts_per_capita, lag.max = 10, na.action = na.pass)  # Adjust the lag.max as needed
```

```{r}

energy_aid_indicators_renewable_watts_relevant_topics_max <- energy_aid_indicators_renewable_watts_relevant_topics %>% 
  group_by(recipient_name) %>% 
  slice(which.max(disbursement))


renewable_topics_plot <-  ggplot(data = energy_aid_indicators_renewable_watts_relevant_topics, aes(x = year, y = Renewable_watts_per_capita, color = topic, label = recipient_name, size = disbursement)) +
    geom_point() +
  theme_minimal()

ggplotly(renewable_topics_plot)
```

```{r}
# Load the necessary packages if not already loaded
library(ggplot2)

# Calculate the maximum values for each topic while preserving "year" and "region" columns
max_values <- aggregate(data = energy_aid_indicators_renewable_watts_relevant_topics, 
                        Renewable_watts_per_capita ~ topic + year + recipient_name + sector + disbursement + sum_disbursement, FUN = max)

# Create the ggplot visualization using max_values
ggplot(data = max_values, aes(x = year, y = Renewable_watts_per_capita, color = topic)) +
  geom_line() +
  geom_point() +
  theme_minimal()


```


