---
title: "Education Aid Indicator"
author: "William Kim"
date: "9/14/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(plotly)
library(readxl)
library(janitor)
library(sunburstR)
library(htmltools)
library(dplyr)
library(shiny)
options(scipen = 99)
```


Import Data from "Monitoring GLobal Aid with Machine Learning"

```{r, include=FALSE, results='hide'}
activity_clusters <- read_csv(here("data/activity_cluster_data.csv"))
```


```{r, include=FALSE, results='hide'}
descriptive_stats <- read_excel(here("data/descriptive_statistics_clusters.xlsx"))
```

Import Data from SIDS Data Platform - *may normalize for population*

```{r}
percent_of_teachers_minimum_qualifications <- read_csv(here("education/sids_app_perecent_teachers_minimum_qualifications.csv"), skip = 1)
```


Replace "No Data with NA"

```{r}
percent_of_teachers_minimum_qualifications[percent_of_teachers_minimum_qualifications == 'No Data'] <- NA
```

Pivot Wider to have Countries in recipient_name column instead of as individual columns

```{r}
percent_of_teachers_minimum_qualifications_longer <- percent_of_teachers_minimum_qualifications %>% 
  mutate(across(!Year, as.character)) %>%
  pivot_longer(cols = !Year,
               names_to = "recipient_name",
               values_to = "value") %>% 
  clean_names()
```




*Filter and Join Data from "Monitoring Global Aid with Machine Learning"*


Calculate Disbursement
- sum_disbursement is all of the aid $ for all topics in a sector for a given year/country
- disbursement is specific to topics 
- For a country/year sum disbursement should equal all of the disbursements under that sector

```{r, include=FALSE, results='hide'}
activity_clusters <- activity_clusters %>% 
  mutate(disbursement = USD_Disbursement * 1000000)
```


Create sids list to filter 


```{r, include=FALSE, results='hide'}
sids <- c('Anguilla', 'Antigua and Barbuda', 'Bahrain', 'Barbados', 'Belize', 'Cabo Verde', 'Comoros', 
             'Cook Islands', 'Dominica', 'Dominican Republic', 'Fiji', 'Grenada', 'Guinea-Bissau', 'Guyana', 
             'Haiti', 'Jamaica', 'Kiribati', 'Maldives', 'Marshall Islands', 'Mauritius', 'Micronesia', 
             'Montserrat', 'Nauru', 'Niue', 'Palau', 'Papua New Guinea', 'Saint Kitts and Nevis', 'Saint Lucia', 
             'Saint Vincent and the Grenadines', 'Samoa', 'Sao Tome and Principe', 'Seychelles', 'Solomon Islands', 
             'Suriname', 'Timor-Leste', 'Tokelau', 'Tonga', 'Trinidad and Tobago', 'Turks and Caicos Islands', 
             'Tuvalu', 'Vanuatu')
```

```{r, include=FALSE, results='hide'}
colnames(activity_clusters)[colnames(activity_clusters) == "cluster"] ="Cluster"
```


Join Data to get topic/sector/disbursements 

```{r, include=FALSE, results='hide'}
activity_stats_join<- full_join(activity_clusters, descriptive_stats, by='Cluster')
```


Filter for SIDS

```{r, include=FALSE, results='hide'}
activity_stats_sids<-activity_stats_join %>%
  filter(RecipientName %in% sids)
```


Clean Column Names

```{r, include=FALSE, results='hide'}
activity_stats_sids <- activity_stats_sids %>% 
  clean_names()
```


Filter for Education Sector Aid

```{r}
activity_stats_sids_education <- activity_stats_sids %>% 
  filter(sector == "Education")
```

Drop columns that are not needed for analysis

```{r}
activity_stats_sids_education <- activity_stats_sids_education %>% 
  select(-cluster, -usd_disbursement, -country_code, -incomegroup_name, -matching_purpose_code, -count_activities, -purpose_code, -area, -sum_disbursement)
```

Change column name so other indicators may be added to dataframe

```{r}
colnames(percent_of_teachers_minimum_qualifications_longer)[colnames(percent_of_teachers_minimum_qualifications_longer) == "value"] ="percent_of_teachers_min_qual"
```

Creatively Join

```{r}
indicator_aid <- activity_stats_sids_education %>%
  left_join(percent_of_teachers_minimum_qualifications_longer, 
            by = c("year", "recipient_name"))
```



*Add other Indicators*

Import Data from SIDS Data Platform 

```{r}
percent_of_teachers_minimum_qualifications <- read_csv(here("education/indicator/sids_app_development-indicators_sdg-SE_TRA_GRDL-BOTHSEX_recentValue_series.csv"), skip = 1)
```